\cleardoublepage
\thispagestyle{empty}

\begin{flushright}
  \LARGE\bfseries Affirmation
\end{flushright}
\medskip

\begin{enumerate}[i. ]
\item Hereby I affirm that I wrote the present thesis without any inadmissible help by a
  third party and without using any other means than indicated. Thoughts that were taken
  directly or indirectly from other sources are indicated as such. This thesis has not
  been presented to any other examination board in this or a similar form, neither in this
  nor in any other country.

\item The present thesis has been prepared since September 2009 at the Institute of
  Algebra, Department of Mathematics, Faculty of Science, TU Dresden under the supervision
  of Prof. Bernhard Ganter.

\item There have been no prior attempts to obtain a PhD at any university.

\item I accept the requirements for obtaining a PhD (Promotionsordnung) of the Faculty of
  Science of the TU Dresden, issued February 23, 2011 with the changes in effect since June
  15, 2011.
\end{enumerate}

\selectlanguage{german}

\begin{flushright}
  \LARGE\bfseries Versicherung
\end{flushright}
\medskip

\begin{enumerate}[i. ]
\item Hiermit versichere ich, dass ich die vorliegende Arbeit ohne unzulässige Hilfe
  Dritter und ohne Benutzung anderer als der angegebenen Hilfsmittel angefertigt habe; die
  aus fremden Quellen direkt oder indirekt übernommenen Gedanken sind als solche kenntlich
  gemacht. Die Arbeit wurde bisher weder im Inland noch im Ausland in gleicher oder
  ähnlicher Form einer anderen Prüfungsbehörde vorgelegt.

\item Die vorliegende Dissertation wurde seit September 2009 am Institut für Algebra,
  Fachrichtung Mathematik, Fakultät Mathematik und Naturwissenschaften, Technische
  Universität Dresden unter der Betreuung von Prof. Bernhard Ganter angefertigt.

\item Es wurden zuvor keine Promotionsvorhaben unternommen.

\item Ich erkenne die Promotionsordnung der Fakultät Mathematik und Naturwissenschaften
  der TU Dresden vom 23. Februar 2011, in der geänderten Fassung mit Gültigkeit vom
  15. Juni 2011 an.
\end{enumerate}

\selectlanguage{english}

\vfill

Dresden, \today

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../main"
%%% End: 
